---
title: "Benfords"
author: "A+T"
date: "2023-01-18"
output: html_document
---

Benford’s Law, or the first-digit law, refers to the fact that many sets of real-life numerical data probabilistically conform to certain patterns. Essentially, it observes that the leading digit of entries is disproportionately likely to be small, with 1 appearing as the first digit roughly 30% of the time. This law applies over a wide variety of data sets, including building heights, street addresses, prices, and population numbers. It has been used to identify fraud, as artificially created data - say, for fake customer lists, or for falsifying elections - will often not conform to Benford’s Law. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
ben=read.csv("benfords.csv")
```


```{r}
library(ggplot2)
base= ggplot(data = ben,
             aes(x = digit,
                 y = perc))
plot1 = base + geom_bar(fill ="purple",
                        stat = 'identity') 
plot1
```

```{r}
titleText='Expected Frequency of Each Digit Appearing First in Data'
subtext='How often will each digit be the leading digit according to Benford?'
sourceText='Source: Benfords Law'
plot2 = plot1 + labs(title=titleText,subtitle = subtext,
                     x = 'First Digit',
                     y = 'Frequency',
                     caption = sourceText)
plot2
```
```{r}
plot3 = plot2 ##+ geom_hline(yintercept = 11.11, #where
                      ##   linetype="dashed",
                        ## size=1.5, #thickness
                        ## alpha=0.1) #transparency
plot3
```
```{r}
library(scales) # for "unit_format""

# customize Y axis
plot4 = plot3 + scale_y_continuous(breaks=c(0,5,10,15,20,25,30,35),
                                   limits = c(0, 35),
                                   labels=unit_format(suffix = '%'))

plot4
```

```{r}
paste0(round(ben$perc,2), '%')

LABELS=paste0(round(ben$perc,2), '%')
###
plot5 = plot4 + geom_text(vjust=0, #hjust if flipping
                          size = 3,
                          aes(y = perc ,
                              label = LABELS))

plot5
```

```{r}
ben$digit <- as.factor(ben$digit)
```

```{r}
plot6 = plot5 + theme_classic()

plot6
```

```{r}
plot7 = plot6 + theme(
  panel.background = element_rect(fill = "lightyellow")
  )

plot7
```
```{r}
plot8 = plot7 + theme(plot.caption = element_text(hjust = 0),
                      plot.title = element_text(hjust = 0.5),
                      plot.subtitle = element_text(hjust = 0.5))
plot8
```

